\Title

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

In fact, it is possible to adapt the HD-SNN model specifically for common computer vision tasks. First, neural addresses are defined to represent the range of possible positions and polarities. In order to simulate such event-based computations on standard CPU- or GPU-based computers and to benefit from parallel computing acceleration, we transform this event-based representation into a dense representation. Indeed, by discretizing time, we transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$. The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$. Similarly, one may discretize the connectivity of the HD-SNN model defined above. The longest synaptic delay defines the depth $\Ktime$ of the kernel, so that all possible delays associated with the different presynaptic addresses are represented. In particular, for each class $\class$ of the supervision task, the entire synaptic set can be represented as a kernel, which is represented by the dense matrix $\kernel$ of size $(\Nclass, \Npol, \Ktime, \Kx, \Ky)$, where $\Nclass$ is the number of classes, $\Ktime$ the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions. To keep the analogy with the HD-SNN model, $\kernel$ gives the weights as a function of class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$. In our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 12 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (36, 2, 21, 17, 17)$ (that is, $436968$ parameters). Such a kernel defines a dense representation of the full synaptic set.

Then, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:stream_b} corresponds to a weighted reordering of the input $A$ with each kernels and positions assigned to the postsynaptic neurons~\citep{grimaldi_learning_2022}. Let's define \emph{evidence} as the logit of a probability, that is, the inverse sigmoid of that probability. By this definition of evidence, logistic regression takes advantage of the fact that if different independent observations (here, the estimated motion at different spatial locations and timings) share a common cause (here, the rigid local motion of the image on the receptive field), then an optimal estimate of the evidence of this motion is the sum of the evidences from the independent sources. Interpreting the weights of the kernel as evidences (also called factors in logistic regression), we may therefore define the activity $B$ of post-synaptic neurons as the integration of this activity in each voxel and for each channel $\class$ in order to infer the evidence of each motion:
%
\begin{linenomath*}
\begin{equation}\label{eq:kernel_b}
\forall x, y, \timev, \; \;
B(\class, x, y, \timev)
= \sum_{p, \delta_x, \delta_y, \synapticdelay_\timev} \kernel(\class, p, \delta_x, \delta_y, \synapticdelay_\timev) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay_\timev)
\end{equation}
\end{linenomath*}
%
where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses within a kernel and $\delta_\timev$ is the synaptic delay. In this formulation, we recognize that it takes advantage of the position invariance observed in images and exploited in convolutional neural networks. Here, we further assume that synaptic motifs should be similar across different times as defined in the temporal convolution. As a consequence, this defines a 3D, spatiotemporal convolutional operator, in which layer of neurons assigned to specific kernels form channels. Using this dense representation, the model's processing of the input $A$ can be written as layer-wise convolution similar to that used in CNNs: $B = \kernel \ast A$ (\seeFig{model} for an illustration). A similar type of spatiotemporal filtering is used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}. In addition,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm that implements a motion estimation task. By assuming locally a constant motion, the authors assume that the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the motion.  For convenience, the connectivity of the neuron $\postsynaddr$ is defined locally around its position $(x_\postsynaddr, y_\postsynaddr)$. Note also that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information gives an answer at the present time that is, after a time equal to the depth $\Ktime$ of the kernel. This well-known computation defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to learn the classification of different motifs in the event stream. % Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.

Keeping the analogy with spiking neurons, the analog activity $B$ represents the integration of synaptic activity, and we will now try to define the detection of spiking motifs using the spatiotemporal kernels. The 3D convolution represents the linear integration of the spike train as the linear input to the neuron. It represents the evidence accumulated from a set of sources in logistic regression. Since we know that at each instant, there is one and only one motion, we will define the activation function of our model as a softmax function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, and analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}. In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a softmax function of the linear combination of the list of events. Formally, using the kernels, the input raster plot is transformed into a probability with the following formula:
%
% TODO : change to logistic regression
\begin{linenomath*}
\begin{equation}\label{eq:mlr}
\forall x, y, \timev, \forall \class \in [1, \Nclass], \; \;
Pr(k=\class \; \vert \; x, y, \timev) =
\frac {\exp (B_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
\end{equation} 
\end{linenomath*}
%
where $Z(x, y, \timev)=\sum_{\class \in [1, \Nclass]} \exp (B_\class(x, y, \timev) +\bias_\class)$ is the partition function, and $\bias_\class$ is a scalar representing the bias associated with the class $\class$. 
In particular, we expect that some specific motifs may generate tightly synchronized outputs as they are integrated in the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike. The spiking output of the model corresponds for instance to the event with the highest probability.

>>>>

In fact, it is possible to adapt the HD-SNN model specifically for common computer vision tasks. First, neural addresses are defined to represent the range of possible positions and polarities. In order to simulate such event-based computations on standard CPU- or GPU-based computers and to benefit from parallel computing acceleration, we transform this event-based representation into a dense representation. Indeed, by discretizing time, we transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$. The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$. Similarly, one may discretize the connectivity of the HD-SNN model defined above. The longest synaptic delay defines the depth $\Ktime$ of the kernel, so that all possible delays associated with the different presynaptic addresses are represented. In particular, for each class $\class$ of the supervision task, the entire synaptic set can be represented as a kernel, which is represented by the dense matrix $\kernel$ of size $(\Nclass, \Npol, \Ktime, \Kx, \Ky)$, where $\Nclass$ is the number of classes, $\Ktime$ the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions. To keep the analogy with the HD-SNN model, $\kernel$ gives the weights as a function of class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \synapticdelay_\timev, \delta_x, \delta_y) = \synapticweight$. In our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 12 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (36, 2, 21, 17, 17)$ (that is, $436968$ parameters). Such a kernel defines a dense representation of the full synaptic set.

Then, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:stream_b} corresponds to a weighted reordering of the input $A$ with each kernels and positions assigned to the postsynaptic neurons~\citep{grimaldi_learning_2022}. Let's define \emph{evidence} as the logit of a probability, that is, the inverse sigmoid of that probability. By this definition of evidence, logistic regression takes advantage of the fact that if different independent observations (here, the estimated motion at different spatial locations and timings) share a common cause (here, the rigid local motion of the image on the receptive field), then an optimal estimate of the evidence of this motion is the sum of the evidences from the independent sources. Interpreting the weights of the kernel as evidences (also called factors in logistic regression), we may therefore define the activity $B$ of post-synaptic neurons as the integration of this activity in each voxel and for each channel $\class$ in order to infer the evidence of each motion:
%
\begin{linenomath*}
\begin{equation}\label{eq:kernel_b}
\forall x, y, \timev, \; \;
B(\class, \timev, x, y)
= \sum_{p, \synapticdelay_\timev, \delta_x, \delta_y} \kernel(\class, p, \synapticdelay_\timev, \delta_x, \delta_y) \cdot A(p, \timev-\synapticdelay_\timev, x - \delta_x, y - \delta_y)
\end{equation}
\end{linenomath*}
%
where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses within a kernel and $\delta_\timev$ is the synaptic delay. In this formulation, we recognize that it takes advantage of the position invariance observed in images and exploited in convolutional neural networks. Here, we further assume that synaptic motifs should be similar across different times as defined in the temporal convolution. As a consequence, this defines a 3D, spatiotemporal convolutional operator, in which the layers of neurons assigned to specific kernels form channels. Using this dense representation, the model's processing of the input $A$ can be written as layer-wise convolution similar to that used in CNNs: $B = \kernel \ast A$ (\seeFig{model} for an illustration). A similar type of spatiotemporal filtering was used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}. In addition,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm that implements a motion estimation task. By assuming locally a constant motion, the authors assume that the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the motion.  For convenience, the connectivity of the neuron $\postsynaddr$ is defined locally around its position $(x_\postsynaddr, y_\postsynaddr)$. Note also that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information provides an inference at the present time that is, after a time equal to the depth $\Ktime$ of the kernel. This well-known computation defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to learn the classification of different motifs in the event stream. % Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation. 
Such a method contrasts with classical methods for delay learning which explictly manipulate the delay as a variable and which are not directly diferentiable~\citep{nadafian_bio-plausible_2020}.

Keeping the analogy with spiking neurons, the analog activity $B$ represents the integration of synaptic activity, and we will now try to define the detection of spiking motifs using the spatiotemporal kernels. The 3D convolution represents the linear integration of the spike train as the linear input to the neuron. It represents the evidence accumulated from a set of sources in logistic regression. Since we know that at each instant, there is one and only one motion, we will define the activation function of our model as a softmax function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, and analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}. In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a sigmoid function $\sigma(\beta) = \frac {1} {1 + \exp(-\beta)}$ of the linear combination of the list of events. Formally, using the kernels, the input raster plot is transformed into a probability with the following formula:
%
% TODO : change to logistic regression
% \begin{linenomath*}
\begin{equation}\label{eq:mlr}
\forall x, y, \timev, \forall \class \in [1, \Nclass], \; \;
Pr(k=\class \; \vert \; x, y, \timev) =
\sigma (B(\class, \timev, x, y) +\bias_\class) 
\end{equation} 
% \end{linenomath*}
%
where $\bias_\class$ is a scalar representing the bias associated with the class $\class$. 
In particular, we expect that some specific motifs may generate tightly synchronized outputs as they are integrated in the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike. The spiking output of the model corresponds for instance to the event with the highest probability.
